---
title: "Data Exploration & Deterministic Functions"
author: "Ednita Tavarez-Jimenez" 
Date: "October 1, 2022"
output: 
  html_document: 
  toc: True
  toc_float: True
  theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1: Terrain Histograms 

Create histograms for the three terrain variables: elevation, slope, and aspect.
Plot all three histograms in one figure and include it in your report.


```{r}
require(here)
here()
dat_hab <- read.csv(here("data/hab.sta.csv"))

par(mfrow = c(1, 3))
hist(dat_hab$elev,col = "pink", main = "Dat Hab & Elevation", xlab = "Elevation (m)", breaks = 25)
hist(dat_hab$slope, col = "light yellow", main = "Dat Hab & Slope", xlab = "Slope (%)", breaks = 25)
hist(dat_hab$aspect, col = "sky blue",border = "black", main = "Dat Hab & Aspect Ratio", xlab = "Aspect (degrees)", breaks = 10)
```

## Q2: Elevation Histogram Interpretation

The elevation graph above you can see ranges from 0-1000 on the x-axis and 0-150 on the y-axis. In addition, we can see that it has a few peaks. There are a few tall peaks with the high frequencies, however, when we look beyond the peak we can see that there are many low elevation sampling sights. The graph is skewed to the right. The frequency drops once the elevation gets past 500. 

## Q3: Slope Units

What are the units of slope in this data set?

elevation (m), percent slope (%), aspect (degrees) 

## Q4: Slope Histogram Interpretation

Consider the distribution of slopes at the bird census sample sites.

I don't see any sampling's that are flat in the slope graph. However, I do see a mixture of steep and shallow slopes. Once the slope becomes more than about 70% steep the frequency begins to drop. Its interesting to see that at the lowest and highest slow there was not a high number of frequency.

## Q5: Aspects 

Briefly define aspect, describing the units used in this dataset

Aspect is the proportional relationship between width and height of the data's plot axis. The units used in this database is degrees. 

## Q6: Aspect Histogram Interpretation

I do not think the sampling sites tend to be north facing. I think they are Southeast facing and the graph is not evenly distributed. In fact, seems to be binomal or random.

## Q7: Terrain/Basal Area Linear Models

Create scatterplots of total basal area and each of the the terrain variables: elevation, slope, and aspect.
Basal area should be on the y-axis.
Visually inspect the plots and fit a linear function to each terrain variable.
Review the linear model parameterization section of the assignment walkthrough if needed.

```{r pressure, echo=FALSE}

line_point_slope = function(x, x1, y1, slope)
{
  get_y_intercept = 
    function(x1, y1, slope) 
      return(-(x1 * slope) + y1)
  
  linear = 
    function(x, yint, slope) 
      return(yint + x * slope)
  
  return(linear(x, get_y_intercept(x1, y1, slope), slope))
}
par(mfrow = c(3, 1))

plot(
  x = dat_hab$elev,
  y = dat_hab$ba.tot,
  xlab = "Elev", 
  ylab = "Total Basal Area", 
  main = "Total Basal Area & Elev", 
  col = "pink")
curve(line_point_slope(x, x1 = 5, y1 = 5, slope = .05), add = TRUE)

plot(
  x = dat_hab$slope, 
  y = dat_hab$ba.tot,
  xlab = "Slope",
  ylab = "Total Basal Area",
  main = "Total Basal Area & Slope", 
  col = "blue",
  cex = .1
)
curve(line_point_slope(x, x1 = 5, y1 = 5, slope = .5), add = TRUE)

plot(
  x = dat_hab$aspect, 
  y = dat_hab$ba.tot,
  xlab = "Aspect",
  ylab = "Total Basal Area",
  main = "Total Basal Area & Aspect", 
  col = "black",
  cex = .3
)
curve(line_point_slope(x, x1 = 5, y1 = 5, slope = .2), add = TRUE, col = "red")

```

## Q8: Terrain/Basal Model Interpretation

For each terrain variable (elevation, slope, aspect), describe the relationship you observe and your model fit. You should consider

**Is there a noticeable association? The noticeable association I see is that they all have x1 & y1 = 5.  Otherwise I don't really see an association between them.**


If so, is it linear? Based on a visual assessment, is your linear model a good fit for the data, why or why not? 

**I don't think that the the linear model is a good fit because there are so many data points and its makes it hard to see.**

