---
title: "Week 6: Reading Questions"
author: "Ednita J Tavarez-Jimenez"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1: In a short paragraph, describe a baseline scenario regarding seed predation. At the end, state the null hypothesis for seed predation.

Baseline scenarion: Looking at the species Polyscias fulva and Pseudospondias microcarpa, the scenario is looking into if any seeds were or were not taken. I want to know about which seeds are eatened or uneatened the larger or the smaller. I predict that the larger seeds will eat be taken than the larger seeds.


Null hypothesis: Smaller seeds were less eaten the the larger seeds therefore, the null hypthesis will be rejected. 



## Q2: Paste the R code you used to complete the table and calculate the rates.
```{r seeds}
rm(list = ls())

pol_n_predation = 26
pol_n_no_predation = 184
pol_n_total = 210
pol_predation_rate = (pol_n_predation / pol_n_total)

psd_n_predation = 25
psd_n_no_predation = 706
psd_n_total =731
psd_predation_rate = (psd_n_predation / psd_n_total)
  
predation_rate <- pol_predation_rate / psd_predation_rate

print(predation_rate)

print(
  paste0(
    "The seed predation rate for Polyscias fulva is: ",
    round(pol_predation_rate, digits = 3))) 

print(
  paste0(
    "The seed predation rate for Pseudospondias microcarpa is: ",
    round(psd_predation_rate, digits = 3)))
```

## Q3: Show your table with the missing values filled in.


```{r pressure, table}

psd_predation_rate = matrix(c(1:5), ncol=5, byrow=TRUE)
 
# specify the column names and row names of matrix
colnames(data) = c('Species','Polyscias fulva (pol)','Pseudospondias microcarpa (psd)')
rownames(data) <- c('any taken','None Taken','N','Predation')
 
# assign to table
final=as.table(data)
 
# display
final

```

## Q4: Report the seed ratio of seed predation proportions and show the R code you used to do the calculation.