---
title: "R Notebook"
output: html_notebook
---

**Q1: Create separate boxplots of body mass for male and female Adelie penguins. Your boxplots do not have to be in the same panel like mine. Show the R-code you used to make the plots.**

```{r}
require(palmerpenguins)
dat_ade = droplevels(subset(penguins, species == "Adelie"))
hist(dat_ade$body_mass_g, main = "Adelie Penguins: Body Mass by sex", xlab = "body mass (g)")

boxplot(dat_ade$body_mass_g ~ dat_ade$sex, main = "Adelie Penguins: Body Mass by sex", xlab = "Sex", ylab = "Body Mass (g)")

dat_ade_f <- droplevels(subset(dat_ade, sex == "female"))

dat_ade_m <- droplevels(subset(dat_ade, sex == "male"))
```
**Q2: Perform a one-sample t-test of the alternative hypothesis that female Adelie penguins have a body mass different from zero grams. Note that this is a very silly alternative hypothesis. Is this a one- or two-tailed test? Show your R-code.**

This is a two-tailed t-test since we are 

```{r}

t.test(dat_ade_f$body_mass_g, data = dat_ade_f, alternative = "less")
```
**Q3: Describe your conclusions based on the p-value of the t-test.**


**Q4 (2 pts.): Now, conduct a slightly less silly test: perform a one-sample t-test of the alternative hypothesis that male Adelie penguins have a mean body mass greater than 4000 grams. Is this a one- or two-tailed test?**

```{r}
t.test(dat_ade_m$body_mass_g, y= NULL, data = dat_ade_m, mu = 4000, alternative = "greater")
```

**Q8: 

```{r}
t.test(dat_ade_f$body_mass_g, dat_ade_m$body_mass_g, alternative = "greater")
```

```{r}
t.test(dat_ade_f$body_mass_g, dat_ade_m$body_mass_g, alternative = "less")
```

