---
title: "Models 1"
author: "Ednita Tavarez-Jimenez"
Date: "November 11, 2022"
output: 
    html_document:
    toc: true
    toc_float: true
    theme: "journal"
editor_options: 
  chunk_output_type: console
---
## Catastrophic Rate Histogram

**Q1: Create a histogram of the salamander reproduction catastrophic rates. Make sure you include an appropriate title and label for the x-axis.**

```{r}
require(here)
catrate <- read.csv(here("data", "catrate.csv"))
head(catrate)
summary(catrate)

hist(catrate$cat.rate, main = "Histogram of Catastrophe Rates", xlab = "Catastrophe Rate", col = "Pink")
```

## Normality Test

**Q2: Conduct a Shapiro-Wilk test of normality of the salamander catastrophic rates. Report the p-value and show the R-code you used to conduct the test.**

The p-value of the Shapiro-Wilk test of normality is .04097

```{r}
shapiro.test(catrate$cat.rate)
```

**Q3: What is the null hypothesis for the Shapiro test?**

The null hypothesis for the shapiro test is that the data were sampled from a normally-distributed population. 

**Q4: Based on the Shapiro test results, is there strong evidence that the sample came from a non-normally-distributed population?**

The p-value was below the suggested .05 p-value, therefore, the null hypothesis not rejected and there is strong evidence that the sample from a non-normally distributed population. 

##T-Test 1

**Q5: Show the code you used to conduct the t-test.**
```{r}
t.test(catrate$cat.rate, mu = 0.2857143)
```

**Q6: State the null hypothesis of the test, in plain nontechnical English.**

The null hypothesis is that the average of residents, citizens, natives, etc where the data was gathered is not different from x which is a number of data values.

**Q7: Is this a one- or two-tailed test?**

This is a one tailed test. 

##T-Test 2

**Q8: What is the p-value from your t-test? Interpret the p-value as a false-positive rate using nontechnical English that a non-scientist would understand.**

The p-value of my t-test is 1.288e-05. 
```{r}
t.test(catrate$cat.rate - catrate$pond)
```

**Q9: What is the confidence interval for the difference between the null hypothesis and alternative hypothesis means? Did it include zero?**

The confidence interval is -5.163773

**Q10: Considering the results from your t-test, did you conclude that there was strong evidence to reject the null hypothesis?**

Considering the results from my t-test, I conclude that there is strong evidence to reject the null hypothesis. 

## Wilcoxon Test 1

**Q11: Show the code you used to conduct the test.**
```{r}
wilcox.test(catrate$cat.rate, mu = 2 / 7)
```

**Q12: Compare the p-value with the p-value you got from the t-test.**
In the Wilcox test I got a p-value of .006275, and in the t-test my p-value was 1.288e-05, they are different. 

**Q13: Considering the results from your rank sum test, did you conclude that there was strong evidence to reject the null hypothesis?**

I conclude that there is not strong evidence to reduce the null hypothesis. 

##Flippers-Normality Tests

**Q16: Show the R-code you used to conduct tests of normality for the flipper lengths of Chinstrap and Adelie penguins.**
```{r}
require(palmerpenguins)
penguin_dat = droplevels(subset(penguins, species != "Gentoo"))
summary(penguin_dat)

boxplot(
  flipper_length_mm ~ species, 
  data = penguin_dat,
  ylab = "Flipper Length (mm)")

dat_adelie = subset(penguin_dat, species == "Adelie")
dat_chin = subset(penguin_dat, species == "Chinstrap")

shapiro.test(dat_adelie$flipper_length_mm)
shapiro.test(dat_chin$flipper_length_mm)
```

**Q17: Interpret the test results. Do you conclude that the flipper lengths are normally-distributed for each species? Make sure your answers make reference to the test output.**

Since the p-value for both is over .05, I conclude that the flipper lengths are normally distributed.

**Q18: Save your figure to a file and include it in your report. Your figure needs to have appropriate dimensions such that the two histograms are not vertically stretched.**

```{r}
par(mfrow = c(1, 2))
hist(dat_adelie$flipper_length_mm, main = "Adeline Flipper Lengths", xlab = "Flipper Length (mm)", ylim = c(0,55),xlim = c(170, 210), col = "pink")
hist(dat_chin$flipper_length_mm, main = "Chinstrap Flipper Lengths", xlab = "Flipper Length (mm)", ylim = c(0,20), xlim = c(170,220), col = "steelblue")
```

## Flipper T-test

**Q19: State the alternative hypothesis of the test, in plain nontechnical English. Consider whether you used a one- or two- tailed test.**

The alternate hypothesis of the test is that there is true difference in means between groups. A two tailed test was used comparing the two data sets. 

**Q20: Include the code you used to conduct the t-test.**

```{r}
t.test(flipper_length_mm ~ species, data = penguin_dat)
levels(penguin_dat$species)