---
title: "Lab 12: Beyond the General Linear Model"
author: "Ednita Tavarez-Jimenez"
Date: "December 5, 2022"
output: 
    html_document:
    toc: true
    toc_float: true
    theme: "journal"
editor_options: 
  chunk_output_type: inline
---

#Lowess

Instructions

Review the population dispersion and statistical power simulation you created for Lab 11.
Re-run the simulation with with a small number of replicates, say 30, so that you get a jagged plot.
Fit a LOWESS model using 30% of the data.
Create a plot like the one I showed above of your simulated data and your fitted LOWESS model. Your plot must show:
The smooth LOWESS line
Points representing the simulation data.
An appropriate legend, title, and axis labels

**Q1: Include your plot in your lab report.**

```{r}
require(here)
dat_bird <- read.csv(here("data", "bird.sub.csv"))
dat_hab <- read.csv(here("data", "hab.sub.csv"))
birdhab <- merge(dat_bird, dat_hab)
dim(birdhab)

fit_1 <- lm(birdhab$BRCR ~ birdhab$ls, data = birdhab)
```

```{r}
#Build Simulation

fit_1_coefs = coefficients(fit_1)
str(fit_1_coefs)

fit_1_summary = summary(fit_1)
str(fit_1_summary)
fit_1_summary$sigma

int_obs <- fit_1_coefs[1]
slope_obs <- fit_1_coefs [2]
sd_obs <- fit_1_summary$sigma
```

```{r}
alpha = 0.05
n_sims = 30
p_vals = numeric(n_sims)

sample_sizes = seq(5, 20)
sample_size_powers = numeric(length(sample_sizes))

# The maximum x value in the simulation.
# Use the maximum observed x-value in the data
max_x = max(birdhab$ls)

linear = function(x, y_int, slope)
{
  return(x*slope + y_int)
  
}

linear(x = 1, y_int = 1, slope = 1)
linear(x = 3:5, y_int = 1, slope = 1)
linear(x = 3:5, y_int = -1, slope = 1)
linear(x = 3:5, y_int = -1, slope = 0.01)

linear_simulator = function (x, y_int, slope, st_dev)
{
 
  int_1 <- linear(x, y_int, slope)
  noise <- rnorm(n = length(x), mean = 0, sd = st_dev)
  noisy_line <- int_1 + noise
  
  return(noisy_line)
  
}

linear_sim_fit = function(x, slope, y_int, st_dev)
{
  y_sim = linear_simulator(
    x = x,
    y_int = y_int,
    slope = slope,
    st_dev = st_dev
  )
  fit_sim = lm(y_sim ~ x)
  return(fit_sim)
}


for(j in 1:length(sample_sizes))
{
  # A sequence of equally-spaced x-values:
  x_vals = seq(0, max_x, length.out = sample_sizes[j])
  
  for(i in 1:n_sims)
  {
    fit_sim = linear_sim_fit(
      x = x_vals,
      y_int = int_obs,
      slope = slope_obs,
      st_dev = sd_obs
    )
    p_vals[i] = summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
  }
  sample_size_powers[j] = sum(p_vals < alpha) / n_sims
}


sim_sample_size = 
  data.frame(
    sample_size = sample_sizes,
    power       = sample_size_powers)

plot(
  power ~ sample_size, data = sim_sample_size,
  type = 'l', xlab = 'Sample size', ylab = 'Power')

```

```{r}
fit_lowess_50 = loess(power ~ sample_size, data = sim_sample_size, span = 0.5)
```

```{r}
newdata_sample_size = data.frame(sample_size = seq(2, 20, length.out = 100)) 
```

```{r}
plot(
  power ~ sample_size, data = sim_sample_size,
  type = 'p', xlab = 'Sample size', ylab = 'Power', col = "red", main = "Ednita's LOWESS Graph")

points(
  x = newdata_sample_size$sample_size,
  y = predict(fit_lowess_50, newdata = newdata_sample_size),
  type = "l", add = TRUE,
  ylab = "Statistical Power", xlab = "Sample Size", lty = "solid")
```

# NLS

**Q2: Include your plot in your lab report.**

Instructions

Follow the lab walkthrough and make sure you can recreate the plots.
Review the exponential model you fit in lab 5.
Fit an NLS exponential model to the marbled salamander data.
Create a plot of your NLS and visually-fitted exponential models. Your plot must include:
A line for your visually-estimated model.
A line for your model fit via NLS.
Appropriate legend, title, and axis labels.

```{r}

require(here)
here()
dat_dispersal = read.csv(
  here("data", "dispersal.csv"))

line_point_slope = function(x, x1, y1, slope)
  {
    get_y_intercept = 
      function(x1, y1, slope) 
        return(-(x1 * slope) + y1)
    
    linear = 
      function(x, yint, slope) 
        return(yint + x * slope)
    
    return(linear(x, get_y_intercept(x1, y1, slope), slope))
}

plot(
    dat_dispersal$dist.class,
    dat_dispersal$disp.rate.ftb,
    xlim = c(0, 1500),
    xlab = "distance class", 
    ylab = "standardized dispersal rate", 
    main = "Ednita's Marbled Salamander Plot - first time breeders")

ricker_fun = function(x, a, b) 
{
  return(a * x * exp(-b * x))
}

exp_fun = function(x, a, b) 
  
{
  return(a * exp(-b * x))
}


fit_ricker_nls = nls(
  disp.rate.ftb ~ ricker_fun(dist.class, a, b),
  data = dat_dispersal,
  start = list(b = 0, a = 1))
summary(fit_ricker_nls)

#first step is to make a data frame

dist_newdata = data.frame(dist.class = seq(0, 1600, length.out = 1600))  


#plot the results of our model

lines(predict(fit_ricker_nls, newdata = dist_newdata))
legend(
  "topleft",
  legend = c("nls fit", "guess"),
  pch = 16,
  col = c(1, adjustcolor("red", alpha = 0.3)))


curve(ricker_fun(x, .005, 1/300), 
  from = 0, to = 1500, add = TRUE, ylim = c(0,1500),
  ylab = "f(x)", xlab = "x",col = "pink", lty = "dotted")

```

# Logistic Models 1: Model Fits

**Q3: What are the AIC values for each of the 4 models?**



```{r}
dat_bird = read.csv(here("data", "bird.sta.csv"))
dat_habitat = read.csv(here("data", "hab.sta.csv"))
dat_all = merge(dat_bird, dat_habitat)

dat_all$HEWA_pres = dat_all$HEWA > 0

# Hermit warbler presence/absence
dat_all$HEWA_pres = dat_all$HEWA > 0

# Create model fits
fit_hewa_slope = glm(HEWA_pres ~ slope, data = dat_all, family = binomial)
fit_hewa_ba_tot = glm(HEWA_pres ~ ba.tot, data = dat_all, family = binomial)
fit_hewa_both_additive = glm(HEWA_pres ~ slope + ba.tot, data = dat_all, family = binomial)
fit_hewa_both_interactive = glm(HEWA_pres ~ slope * ba.tot, data = dat_all, family = binomial)

summary(fit_hewa_both_additive)

n = 500

slope_newdata = data.frame(
  slope = seq(
    from = min(dat_all$slope, na.rm = T),
    to = max(dat_all$slope, na.rm = T),
    length.out = n
  )
)

#new data.frame of total basal area values for prediction. Call it ba_newdata.

ba_newdata = data.frame(
  ba.tot = seq(
    from = min(dat_all$ba.tot, na.rm = T),
    to = max(dat_all$ba.tot, na.rm = T),
    length.out = n))


slope_newdata$hewa_predicted = 
  predict(
    fit_hewa_slope,
    newdata = slope_newdata,
    type = "response"
  )

ba_newdata$hewa_predicted = 
  predict(
    fit_hewa_ba_tot,
    newdata = ba_newdata,
    type = "response"
  )

#Compare models using AIC

AIC(
  fit_hewa_ba_tot,
  fit_hewa_slope,
  fit_hewa_both_additive,
  fit_hewa_both_interactive)

```


**Q4: Which model would you choose, and why?**

I would choose the AIC model because it'll show me the best model to use in this case it would be fit_hewa_both_interactive.

```{r}
n = 50

ba.tot = seq(
  from = min(dat_all$ba.tot, na.rm = T),
  to = max(dat_all$ba.tot, na.rm = T),
  length.out = n)
slope = seq(
  from = min(dat_all$slope, na.rm = T),
  to = max(dat_all$slope, na.rm = T),
  length.out = n)

new_dat_all = expand.grid(
  ba.tot = ba.tot,
  slope = slope)
head(new_dat_all)

tail(new_dat_all)

new_dat_all$pred_add = predict(
  fit_hewa_both_additive,
  newdata = new_dat_all,
  type = "response")

new_dat_all$pred_int = predict(
  fit_hewa_both_interactive,
  newdata = new_dat_all,
  type = "response")

z_hewa_add = matrix(
  new_dat_all$pred_add,
  nrow = length(ba.tot),
  byrow = FALSE)

z_hewa_int = matrix(
  new_dat_all$pred_int,
  nrow = length(ba.tot),
  byrow = FALSE)

#3D plot

require(rgl)

rgl::persp3d(
  x = ba.tot,
  y = slope,
  z = z_hewa_add,
  col = "steelblue",
  xlab = "Basal Area",
  ylab = "Slope",
  zlab = "Pr(present)",
  alpha = 0.4)
rglwidget()

```


**Q5: Based on the model coefficient table of your chosen model, describe the direction and significance of the relationship(s) of the predictor variable or variables to the binary response. Make sure your answer is in terms of the ecological context.**

Based on the model coefficient table of the AIC, the relationships between the variables in each model are relatively close to one another. It is going in a pretty stable to downward direction. 

# Logistics Models 2: Graphical Exploration

**Q6: Include your two single-predictor model plots in your report.**

```{r}
par(mfrow = c(2, 1))

# Presence/absence data, translucent points:
plot(
  HEWA_pres ~ slope, data = dat_all,
  xlab = "Percent Slope",
  ylab = "HEWA presence/absence",
  pch = 16, cex = 1.5, col = gray(0, 0.2)
)

lines(hewa_predicted ~ slope, data = slope_newdata)

plot(
  HEWA_pres ~ ba.tot, data = dat_all,
  xlab = "Basal Area",
  ylab = "HEWA presence/absence",
  pch = 16, cex = 1.5, col = gray(0, 0.2)
)
```


**Q7 (4 pts.): Include contour plots (or interactive 3D perspective plots) in your report.**

```{r}
par(mfrow = c(1, 2))
contour(
  x = ba.tot, y = slope,
  z = z_hewa_add,
  xlab = "Total Basal Area",
  ylab = "Percent Slope",
  main = "Additive")
contour(
  x = ba.tot,
  y = slope,
  z = z_hewa_int,
  xlab = "Total Basal Area",
  ylab = "Percent Slope",
  main = "Interactive")
```

