---
title: "Lab 03: Data Exploration and Deterministic Functions"
author: "Ednita Tavarez-Jimenez"
Date: "September 28, 2022"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **Basal Area**

Q5 (1 pt.): What is basal area, and how is it measured?

  The Basal Area is the average number of an area. It is measured square feet per acre.

Q2 (2 pts.): Include a figure of your terrain/basal area pairplot.

```{r}
#install.packages("psych")
require(psych)
require(here)
dat_bird = read.csv(
  here("data", "bird.sta.csv"))
head(dat_bird)

dat_hab = read.csv(here("data", "hab.sta.csv"))

head(dat_hab)
data_all <- merge(dat_bird, dat_hab)
pairs.panels(data_all[, c("slope", "aspect", "elev","ba.tot")])
plot(ba.tot ~ elev, data = data_all)


sample(data_all$CEWA, 100)
my_vec = rep(1:3, 5)
my_vec > 1

sample(data_all$CEWA, 100)
cewa_present_absent <- as.numeric(dat_bird$CEWA > 0)

plot(x = dat_hab$ba.tot, y = cewa_present_absent, main = "Basal Area vs. CEWA Presence", xlab = "Hab Basal Area (ft²/ac)", ylab = "CEWA Absence or Presence")
```

# **Bird Species 1 and 2**

Q3/Q5 (1 pt.): Include a figure of your logistic function plot. Your figure must include the name of the bird species, appropriate title, axes, etc.

**Find Graph Below**

Q4/Q6 (3 pts.): Qualitatively describe the bird’s presence/absence patterns in terms of basal area (or your other chosen predictor).

The Swaison Thrust and the Winter Wren both prefer areas with high tree coverage, there was more presence of these birds the in the high tree cover basal area vs the love basal area. When running the elevations they both preferred higher elevation. Although they were absent past est. 860 elevation. I think it did fit the logistic model plot but it was very confusing on how to interpret it. 

```{r}
get_logistic_param_a = function(slope, midpoint)
{
  b = slope / 4
  return (-midpoint * (slope / 4))
}

# Function to calculate the logistic parameter b given the slope
get_logistic_param_b = function(slope)
{
  return (slope / 4)
}


# Calculate the value of the logistic function at x, given the parameters a and b.
logistic = function(x, a, b)
{
  val = exp(a + b * x)
  return(val / (1 + val))
}

# Calculate the value of the logistic function at x, given a slope and midpoint.
logistic_midpoint_slope = function(x, midpoint, slope)
{
  b = get_logistic_param_b(slope)
  a = get_logistic_param_a(slope, midpoint)
  return(logistic(x, a, b))
}

apply(dat_bird[, -c(1, 2, 3)], 2, sum)
sample(data_all$SCOW, 100)
my_vec = rep(1:3, 5)
my_vec > 1

sample(data_all$SWTH, 100)
SWTH_present_absent <- as.numeric(dat_bird$SWTH > 0)
plot(x = dat_hab$ba.tot, y = SWTH_present_absent, main = "Swaison's Thrush Basal Area Presence", xlab = "Basal Area (ft²/ac)", ylab = "Swaison's Thrush" )
curve(logistic_midpoint_slope(x, midpoint = 75, slope = 0.5), add = TRUE)

sample(data_all$WIWR, 100)
WIWR_present_absent <- as.numeric(dat_bird$WIWR > 0)
plot(x = dat_hab$ba.tot, y = WIWR_present_absent, main = "Winter Wren Basal Area Presence", xlab = "Basal Area (ft²/ac)", ylab = "Winter Wren")
curve(logistic_midpoint_slope(x, midpoint = 75, slope = 0.5), add = TRUE)

#Elev
sample(data_all$SWTH, 100)
SWTH_present_absent <- as.numeric(dat_bird$SWTH > 0)
plot(x = dat_hab$elev, y = SWTH_present_absent, main = "Swaison's Thrush Basal Area Presence", xlab = "Elev", ylab = "Swaison's Thrush" )
curve(logistic_midpoint_slope(x, midpoint = 75, slope = 0.5), add = TRUE)

sample(data_all$WIWR, 100)
WIWR_present_absent <- as.numeric(dat_bird$WIWR > 0)
plot(x = dat_hab$elev, y = WIWR_present_absent, main = "Winter Wren Basal Area Presence", xlab = "Elev", ylab = "Winter Wren")
curve(logistic_midpoint_slope(x, midpoint = 75, slope = 0.5), add = TRUE)
```

# **Gray Jays** 

Q7 (1 pt.): How many total number of Gray Jays were observed in all of the sampling sites.

181 Gray Jays were observed in all of the sampling sites. 

Q8 (2 pts.): Show the R code you used to perform the calculation.

sum(data_all$GRJA)

Q9 (1 pt.): Calculate the total number of sampling sites in which Gray Jays were observed.

Gray Jays were observed in 110 sampling sites 

Q10 (2 pts.): Include the R code you used to perform the presence/absence calculation.

sum(GRJA_present_absent)

```{r}
sample(data_all$GRJA, 100)

GRJA_present_absent = as.numeric(dat_bird$GRJA >0)

sum(GRJA_present_absent)

sum(data_all$GRJA)


```

