---
title: "Final R Reference"
author: "Ednita J Tavarez-Jimenez"
Date: "November 21, 2022"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal
editor_options: 
  chunk_output_type: inline

---
# R Reference Guide {.tabset .tabset-pills}


## Loading Data and Packages

```{r}
require(here)
library(palmerpenguins)
require(palmerpenguins)
here()
ginkgo <- read.csv(here("data", "ginkgo_data_2022.csv"))
dat_delo <- read.csv(here("data", "delomys.csv"))
dat_pen <- penguins
```

## Data Structure

```{r}
num_vec  = c(2, 4, 6)

## Create a vector of characters:
char_vec = c("a", "fantastic", "year")

num_vec

print(char_vec)
```
**Length**

```{r}
length(dat_delo)
```

**Matrix**

```{r}
dim(dat_delo)
dat_char <- matrix(nrow = length(num_vec), ncol = length(char_vec))

#data frame for dat_delo
dat_char = data.frame(char_vec, num_vec)
```

## Subsetting

```{r}
ginkgo$seeds_present

ginkgo[1]
ginkgo [2,3]
ginkgo[,3]
Adelie <- subset(penguins, species =="Adelie")
```

## Numerical Data Exploration

```{r}
summary(Adelie)
mean_ad <- mean(Adelie$bill_length_mm, trim = 0, na.rm = TRUE)
sd((!is.na(Adelie$bill_length_mm)))
```

## Graphical Data Exploration 

```{r}
plot(max_depth ~ max_width, data = ginkgo,
     col = "steelblue",
     pch = 11,
     cex = .5,
     main = "Ednita's plot of Ginkgo Max Width and Max Depth",
     xlab = "Max Leaf Depth",
     ylab = "Max Leaf Length", 
     xlim = c(20,120),
     ylim = c(20,150))
     
```
```{r}
hist(dat_pen$flipper_length_mm, 
     col = "pink", 
     main = "Penguin Flipper Length (mm)",
     xlab = "Flipper Length (mm)", 
     xlim = c(160,250),
     ylim = c(0, 60), 
     breaks = 20)
```
```{r}
boxplot(ginkgo$petiole_length, 
        col = "red",
        main = "Ginkgo Pertiole Length",
        ylab = "Petiole Length")
```
```{r}
boxplot(
  max_width ~ seeds_present,
  data = ginkgo,
  col = "Pink",
  las = 2,
  xlab = "Seeds Present",
  ylab = "Max Width",
  main = "Max Width Conditioned on Seeds Present")
```
```{r}
par(mfrow = c(2,2))

hist(dat_delo$body_mass,
     col = "steelblue",
     main = "Delomys Body Mass", 
     xlab = "Delomys Body Mass")
hist(dat_pen$bill_length_mm, 
     col = "pink",
     main = "Penguin's Bill Depth (mm)",
     xlab = "Penguin's Bill Length (mm)")
hist(Adelie$bill_depth_mm,
     col = "Red",
     main = "Adelie Bill Depth (mm)",
     xlab = "Bill Depth (mm)")
hist(dat_delo$body_length, 
     col = "black", 
     main = "Delomys Body Length", 
     xlab = "Body Length")
```

## Distribution Functions

```{r}
dnorm(x = 4, mean = 0, sd = 1)
```

```{r}
pnorm(q = 4, mean = 0 , sd = 1)
```

```{r}
qnorm(.75, mean = 0 , sd = 1)
```

```{r}
dbinom(10, size = 11, prob = .5)

```

```{r}
pbinom(50, size = 51, prob = .75)
```

```{r}
qbinom(.111, size = 11, prob = .5)
```

# Part 2: Data Analysis

# Data Exploration

**Q1: Qualitatively describe the relationship between body mass and length.**

The relationship between body mass and length look a bit linear with a small curve on the scatter plot.

```{r fig.width= 8, fig.height = 8}
require(palmerpenguins)
plot(body_length ~ body_mass, data = dat_delo, main = "Delomys Body Mass and Body Length Histogram", xlab = "Body Mass", ylab = "Body Length", col = "black")
```


**Q2: Qualitatively describe the shapes of the histograms.**

The data does not appear to be normally distributed there is variation in the shape with body mass being a little skewed to the right. We care if the data is normally distributed or not because we would need to decide which model is appropriate to use or not going forward. In addition, we would want to see if there are any outliers if the data is not normally distributed.

```{r fig.width = 10, fig.height = 10}
#Histogram of Penguin Body Mass and Body Length
par(mfrow = c(2, 1))
hist(dat_pen$body_mass_g, main = "Delomys Body Mass Histogram", xlab = "Body Mass", col = "pink")
hist(dat_delo$body_length, xlab = "Body Length",main = "Delomys Penguin Body Length Histogram", col = "steelblue")
```


**Q3 (2 pts.): Using both the histograms and normality tests, do you think the (unconditioned) body masses and body length are normally-distributed?**

Looking at the histogram and normality tests, I do not think the body masses and body length are normally distributed. In the histogram, we can see that the graphs are a bit skewed. In the shapiro test, the p-value is 2e-16 which is a lot less than the 0.05 so the null hypothesis would be rejected.

```{r}
dat_delo_1 = lm(body_mass ~ body_length, data = dat_delo)

summary(dat_delo_1)

shapiro.test(residuals(dat_delo_1))
```


**Q4: Examine the three conditional boxplots.**

Looking at the conditional boxplot I don't see a large variation in body mass between the two species and the sex. In the body mass conditioned by species boxplot we see that the dorsalis weights a little more than the sublineatus. However, in the body mass condition on sex boxplot we can see that the males weight a little more than the females.
```{r fig.width = 10, fig.height = 20}
# Conditional Box Plots

#1. Conditional boxplot of body mass, conditioned on species (column binomial)
par(mfrow = c(3, 1), mar = c(5,3,3,1))
boxplot(body_mass~binomial,data = dat_delo, main = "Delomys Body Mass Conditioned on Species", col = "red", xlab = "Species", ylab = "Body Mass")

boxplot(body_mass~sex,data = dat_delo, main = "Delomys Body Mass Conditioned on Sex", col = "pink", xlab = "Sex", ylab = "Body Mass")

boxplot(body_mass~binomial+sex,data = dat_delo, main = "Delomys Body Mass Conditioned on Species & Sex", col = "steelblue", xlab = "Species", ylab = "Body Mass", names = c("Female\n Delomys\n Dorsalis", "Female\n Delomys\n Sublineatus", "Male\n Delomys\n Dorsalis", "Male\n Delomys\n Sublineatus"), las = 2)
```

# Model Building

```{r fig.width= 8, fig.height = 6}
#Building models for ANOVA
fit1 = 
  lm(
    formula = body_length ~ body_mass,
    data = dat_delo)
coef(fit1)

slope_observed = coef(fit1)

plot(
  body_length ~ body_mass, data = dat_delo,
  main = "",
  xlab = "Body Length",
  ylab = "Body Mass")
abline(fit1)
```


```{r}
fit2  = 
  lm(
    formula = body_mass ~ sex,
    data = dat_delo)
anova(fit2)
```


```{r}
fit3 = 
  lm(
    formula = body_mass ~ binomial,
    data = dat_delo)
anova(fit3)

```


```{r}
fit4 = 
  lm(
    formula = body_mass ~ sex + binomial,
    data = dat_delo)
anova(fit4)
```


```{r}
fit5 = 
  lm(
    formula = body_mass ~ sex * binomial,
    data = dat_delo)
anova(fit5)
```

```{r fig.width = 10, fig.height = 20}
#Histograms of model residuals 
par(mfrow = c(5, 1), mar = c(4,3,3,1))
hist(residuals(fit1), main = "Histogram of residuals fit 1", xlab = "Residuals (fit 1)", col = "pink")
hist(residuals(fit2), main = "Histogram of residuals fit 2", xlab = "Residuals (fit 2)", col = "steelblue")
hist(residuals(fit3), main = "Histogram of residuals fit 3", xlab = "Residuals (fit 3)", col = "red")
hist(residuals(fit4), main = "Histogram of residuals fit 4", xlab = "Residuals (fit 4)")
hist(residuals(fit5), main = "Histogram of residuals fit 5", xlab = "Residuals (fit 5)", col = "green")
```

```{r}
#Shapiro Test fit 1
shapiro.test(residuals(fit1))
```

```{r}
#Shapiro Test fit 2
shapiro.test(residuals(fit2))
```

```{r}
#Shapiro Test fit 3
shapiro.test(residuals(fit3))
```

```{r}
#Shapiro Test fit 4
shapiro.test(residuals(fit4))
```

```{r}
#Shapiro Test fit 5
shapiro.test(residuals(fit5))
```

**Q5: What do you conclude about residual normality based on the numerical and graphical diagnostics?**

Based on the numerical and graphical diagnostics that the residual normality for sex and is not normal. The p-value is well below 0.05 which means the data is not normally distributed. However, based on the numerical and graphical diagnostics on sex and binomial the p-value is over 0.05 which means that the data is normal and normally distributed. 

**Q6: Are violations of the normality assumption equally severe for all the models?**
No it is not equally severe for the sex:binomial which has a p-value of .950442

# Model Interpretation

**Q7: What is the magnitude of the mass/length relationship?**

The magnitude of the mass/length relationship is that it is increasing by .87.

```{r}
knitr::kable(coef(summary(fit1)), digits = 2)
slope_observed = coef(fit1) [2]
summary(coef(fit1))
```

**Q8: What is the expected body length of an animal that weighs 100g?**

The expected body length of an animal that weights 100g is 164.12
```{r}
76.12 + .88*100
```

**Q9: What is the expected body length of an animal that weighs 0g?**

The expected body length for an animal that weighs 0g is 76.12.

```{r}
76.12+.88*0
```


**Q10: What is the base level for sex?**

The base level for sex are the females.
```{r}
knitr::kable(coef(summary(fit2)), digits = 2)
coef(fit2)
summary(fit2)
```


**Q11: What is the base level for binomial?**

Delomys dorsalis is the base level for binomial.
```{r}
knitr::kable(coef(summary(fit3)), digits = 2)
coef(fit3)

summary(fit3)
```


**Q12: Which sex is heavier? How do you know?**

The Male sex is heavier because they weight 45.4841 where as, the females only weight 42.7.
```{r}

female_sex <- 42.7 + 2.7841 * 0 
male_sex <- 42.7 + 2.7841 * 1

```


**Q13: Which species is heavier? How do you know?**

Delomys dorsalis is heavier because they weight 46.7 where as, Delomys sublineatus only weights 39.01.

```{r}
coef(fit3)

summary(fit3)

d_dor <- 46.7 - 7.6831 * 0
d_sub <- 46.7 - 7.6831 *1

```

## Anova Tables

**Q14: Are sex and species significant predictors for body mass?**

Yes because the p-value is under .05 for both the sex and the species in the additive and factorial ANOVA.

```{r}
anova(fit2)
```

```{r}
anova(fit3)
```


```{r}
anova(fit4)
```

```{r}
anova(fit5)
```


**Q15: Is there a significant interaction?**

There not a significant interaction because the p-value is .95 which is above the .05.

**Q16: Examine the p-values for the main effects (sex and species) in all four of the ANOVA tables. Does the significance level of either main effect change very much among the different models?**

Looking at the p-values for the main effects in all four of the ANOVA tables the significance level does not change very much among the different models. 

##Model Comparison

**Q17: Which two models have the lowest AIC?**

The model with the lowest AIC is fit3.
```{r}
AIC(
  fit2,
  fit3,
  fit4,
  fit5)
```


**Q18: Which of the two models with lowest AIC scores would you select?**

I would select fit3 and fit4, due to their low numbers there is strong evidence in favor of the model fit. In addition, they are also the two that are more of frugal which is the trade off. However, these two model could also have a chance of fitting poorly. 


